---
title: "DAS GROUP 8 PROJECT "
author: "Constantinos Anastasiou, Lin Lin, Yuang Tian, Yutong WU, Shengyuan Xia"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

```{r}
#| label: libraries
library(ggplot2)
library(tidyverse)
library(gt)
library(patchwork)
library(gridExtra)
library(moderndive)


data <- read.csv("dataset08.csv")
data
films <- na.omit(data)


films
```

```{r}
#| label: data
data <- read.csv("dataset08.csv")
dataNEW <- na.omit(data)
colnames(dataNEW) <- c("FILM ID", "YEAR RELEASED", "LENGTH", "BUDGET" , "VOTES" ,"GENRE" , "RATING") 

```

# Introduction {#sec-Intro}
The IMDB film database provides a rich source of information on various films. This data set offers valuable insights into the factors influencing a film's critical reception. In response to a film producer's inquiry, we conducted an in-depth analysis using a Generalized Linear Model to investigate the properties of films that influence whether they are rated greater than 7 on IMDB. Through the analysis, we found compelling evidence regarding the impact of different variables. Specifically, we observed that budget, genre, film length, and release year play significant roles in determining a film's reception.

# Exploratory Data Analysis {#sec-EDA}

# Formal Data Analysis {#sec-FDA}

```{r}

#WE CONVERT RATING TO BINARY VARIABLE

films$GREAT <- ifelse(films$rating > 7 , 1 ,0)
films

#FIT GLM
#NOTE WE ARE NOT INCLUDING film id as it obviously doesn't have an effect on the film's rating
glm_model <- glm(GREAT ~ film_id + year + length + budget + votes + genre , data = films , family = binomial)
summary(glm_model)
```

# Conclusion
In conclusion, based on the above analysis, we found that variables such as budget, genre, film length, and release year play crucial roles in determining critical reception. Comedy and documentary genres, lower budgets, shorter film lengths, and more recent release years were associated with higher probabilities of achieving a lower rating. However, our analysis has certain limitations, such as potential confounding variables not included in the model, potential biases in user ratings on IMDB, and the need for further exploration of interaction effects between variables. To improve this analysis, future research could consider incorporating additional variables, exploring nonlinear relationships, and conducting a more in-depth investigation.
